---
title: "QBS120_problemset2"
output:
  pdf_document: default
  html_document: default
  word_document: default
Author: Claire Wang
---

title: "QBS120_problemset2"
Author: Claire Wang


Problem 1 (Based on Rice 3.18):
A. the region here is $y \leqslant x$, with $0 \leqslant x \leqslant 1$

```{r}
plot(1,1, col = "white", xlab = "x", ylab = "y", xlim = c(0,1), ylim = c(0,1))   # Draw empty plot
polygon(x = c(0, 1,1), y = c(0,0, 1), col = "blue", border = NA) # the blue region which shows the limits of the integration where the joint denity function is positive
```

B. find k:
$f(x)=\begin{cases} kxy, 0 \leqslant y \leqslant x \leqslant 1\\0,otherwise\end{cases}$

$\int_{0}^{1} \int_{0}^{x} kxy\ dydx = 1$

$k\int_{0}^{1} [\frac{y^2}{2}] \Big|_0^x\ x\ dx = 1$

$k\int_{0}^{1} \frac{1}{2} x^3 dx = 1$

$k[\frac{x^4}{8}] \Big|_0^1\ = 1$

$\frac{k}{8} = 1$

$k=8$
            
C. find the marginal densities of X and Y:

Marginal density of y:
$f_Y = \int_{y}^{1} 8xy dx = 8 [\frac{1}{2}x^2] \Big|_y^1 y$

$f_Y= 4y(1-y^2)$

Marginal density of x:
$f_X = \int_{0}^{x} 8xy dy = 8 [\frac{1}{2}y^2] \Big|_0^x x$

$f_X= 4x^3$

D. find the conditional densities of Y given X and X given Y:
$f_{Y|X}(y|x) =  \frac{f_{X,Y}(x,y)}{f_X (x)}$

$= \frac{8xy}{4x^3} = \frac{2y}{x^2}$

$f_{X|Y}(x|y) =  \frac{f_{X,Y}(x,y)}{f_Y (y)}$

$= \frac{8xy}{4y(1-y^2)} = \frac{2x}{1-y^2}$








Problem 2 (Based on Rice 3.20):
rephrase: $f_{X_1}$ is a uniform density function with  $x_1 \in [0,1]$, 
          $f_{X_2|{X_1}}$ is a uniform density function with $x_2 \in [x_1, 1]$
          
given that $f_{X_1} = \begin{cases} 1, 0 \leqslant x_1  \leqslant 1 \\0, otherwise\end{cases}$

and $f_{X_2|{X_1}} = \begin{cases} \frac{1}{1-x_1}, x_1 \leqslant x_2  \leqslant 1 \\0, otherwis\end{cases}$

the joint distribution is:

$f_{X_1,X_2}(x_1,x_2) = f_{X_1} \times f_{X_2|X_1} = 1 \times \frac{1}{1-x_1}$ with $x_1 \in [0,1]$ 

and $x_2 \in [x_1, 1]$

the marginal distribution of $X_1$ is given as above.
the marginal distribution of $X_2$ is:

then $f_{X_2} = \int_{0}^{x_2} f_{X_1,X_2} (x_1,x_2) dx_1$

$f_{X_2} = \int_{0}^{x_2} \frac{1}{1-x_1}dx_1=- [\ln(|x_1-1|)]\Big|_0^{x_2} = - [\ln(|x_2-1|)]$
with $0 \leqslant x_1 \leqslant x_2 \leqslant 1$


$f_{X_2}= \begin{cases} - [\ln(|x_2-1|)] , 0\leqslant x_1 \leqslant x_2 \leqslant 1 \\0, otherwise \end{cases}$








Problem 4 (Based on Rice 3.71):
Since $X_1$, $X_2$,..., $X_n$ all have the same density function,
$f_{X_i}(x_i) = f_X(x)$
Given $P(x > \alpha) = \gamma$, find $P(x_1 \leqslant \alpha)$, which is $F_{X_{(1)}}(\alpha)$

now we know $P(x \leqslant \alpha) = F_X(\alpha) = 1 - \gamma$

Given that $F_{X_(n)}(x) = \max\{X_1,X_2,...,X_n\} = [F_X(X)]^n$

and $F_{X_(1)}(x) = \min\{X1,X_2,...,X_n\} = 1-[1-F_X(x)]^n$
      
We have $F_{X_{(1)}}(\alpha) = 1 - [1 - F_X(\alpha)]^n = 1 - [1 - (1-\gamma)]^n = 1 - \gamma^n$








Problem 5 (Based on Rice 4.31):
$f_X = \begin{cases} \frac{1}{2}, 1 \leqslant x \leqslant 3 \\0, otherwise\end{cases}$

$E(\frac{1}{X}) = \int_{1}^{3} \frac{1}{2} \frac{1}{x} dx = \frac{1}{2} [\ln(x)]\Big|_1^3 = \frac{1}{2} \log(3)$

$E(X) = \int_{1}^{3} \frac{1}{2}x dx = [\frac{1}{4} x^2]\Big|_1^3 = 2$

Therefore,

$E(\frac{1}{X}) \neq \frac{1}{E(X)}$




Problem 6 (Based on Rice 4.49):
A.
$E(Z) = E(\alpha X + (1-\alpha)Y) = \alpha E(X) + (1-\alpha)E(Y) = \alpha \mu + (1 - \alpha) \mu = \mu$

B.
$Var(Z) = Var(\alpha X + (1-\alpha)Y) = \alpha ^2 Var(X) + (1- \alpha)^2 Var(Y) = \alpha ^2 \sigma_x^2 + (1-\alpha)^2 \sigma_y^2$
here, $\sigma_x^2$ and $\sigma_y^2$ are constont, so differentiate $Var(Z)$ and set it to equal zero to find the $\alpha$ that minimize $Var(x)$
$[Var(Z)]' = 2\alpha \sigma_x^2 - 2(1- \alpha)\sigma_y^2 = 0$
$\alpha(\sigma_x^2 + \alpha_y^2) - \sigma_y^2 = 0$
$\alpha = \frac{\sigma_y^2 }{\sigma_x^2 + \sigma_y^2}$


C.
if use $Z = \frac{X+Y}{2}$:
$Var(Z) = Var(\frac{X+Y}{2}) = \frac{1}{4}Var(X) + \frac{1}{4}Var(Y) = \frac{1}{4}\sigma_x^2 + \frac{1}{4} \sigma_y^2$
now,
if $Var(\frac{X+Y}{2}) < Var(X)$, $\frac{1}{4}\sigma_x^2 + \frac{1}{4} \sigma_y^2 < \sigma_x^2$, then $\sigma_y^2 < 3\sigma_x^2$
similarly, $Var(\frac{X+Y}{2}) < Var(Y)$, $\frac{1}{4}\sigma_x^2 + \frac{1}{4} \sigma_y^2 < \sigma_y^2$, then $\sigma_x^2 < 3\sigma_y^2$
therefore, 
$\ \frac{1}{3} < \frac{\sigma_x^2}{\sigma^y_2} < 3$







Problem 7 (Based on Rice 5.57):
$X: \mu_x, \sigma_x$
$Y: \mu_y, \sigma_y$
if X and Y are independent RVs,
$E[XY] = E(X)E(Y) = \mu_x \mu_y$
$Var[XY] = E[(XY)^2] - (E[XY])^2 = E(X^2)E(Y^2) - \mu_x^2 \mu_y^2$
since $E(X^2) = Var(X)+[E(X)]^2 = \sigma_x^2 + \mu_x^2$
$Var[XY]= (\sigma_x^2 + \mu_x^2)(\sigma_y^2 + \mu_y^2)- \mu_x^2 \mu_y^2$




Problem 8 
A. The joint empirical distribution of the first two genes in the Golub data set
```{r}
library(multtest)
library(ggpubr)
data("golub")
#View(golub)
golubb <- as.data.frame(golub)
gene <- t(golubb[1:2,])
gene1 <- as.data.frame(gene)
golub.clx <- as.data.frame(golub.cl)
gene1_golub <- cbind(gene1,golub.clx)
colnames(gene1_golub)[1] <- "gene1"
colnames(gene1_golub)[2] <- "gene2"
gene1_golub$ALL_or_AML <- gene1_golub$golub.cl
for (n in 1:38){
  if (gene1_golub[n,3] == "0"){
    gene1_golub[n,4] <- "ALL"
  }
  else{
    gene1_golub[n,4] <- "AML"
  }
}
ggscatter(gene1_golub, "gene1", "gene2", color = "ALL_or_AML",palette = c("purple","yellow"))

```



B.
Since the two genes are positively correlated to each other, they are dependent.

C.
```{r}
plot(density(gene1_golub$gene1))
hist(gene1_golub$gene1)
plot(density(gene1_golub$gene2))
hist(gene1_golub$gene2)
```



D.
they are continuous RV's based on the (multi-modal) distribution plots.


E.
```{r}
# create two sets of 100 random variables and select their mean and standard deviation parameters.
random1 <-rnorm(n = 100, mean(gene1_golub$gene1),sd(gene1_golub$gene1))
random2 <- rnorm(n = 100, mean(gene1_golub$gene2),sd(gene1_golub$gene2))
plot(random1, random2)
#cor(random1, random2)
cor.test(random1,random2)
```
Observing from the scatter plot, the two independent genes we create have no correlation. Also, the p-value is way higher than 0.05, therefore is not statistically significant.

