---
title: "QBS120_P3"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

QBS 120 HW 3

Author: Claire Wang

Corrections will be written in the ** **. 



Problem 1 (Based on Rice 4.54):

(a) Find Cov(U,V) and $\rho_{U,V}$

Given $U = Z - X$ and $V = Z - Y$,
      $Var(X) = \sigma_X^2$, $Var(Y) = \sigma_Y^2$,  $Var(Z) = \sigma_Z^2$

$Cov(Z-X, Z-Y) = Cov(Z,Z-Y) + Cov(-X, Z-Y) = Cov(Z-Y,Z) + Cov(Z-Y,-X)$
$Cov(Z-X, Z-Y) = Cov(Z,Z) + Cov(-Y,Z) + Cov(Z,-X) + Cov(-Y,-X)$

since X, Y, and Z are independent RVs,
$Cov(U,V) = Cov(Z-X, Z-Y) = Cov(Z,Z) = Var(Z)$

Similarly, $\rho_{U,V} = \frac{Cov(Z-X, Z-Y)}{\sqrt{Var(Z-X) Var(Z-Y)}} = \frac{Var(Z)}{\sqrt{[Var(Z)+Var(X)-2\times1\times1\times Cov(Z,-X)]\times[Var(Z)+Var(Y)-2\times1\times1\times Cov(Z,-Y)]}}$

then $\rho_{U,V}=\frac{\sigma_Z^2}{\sqrt{(\sigma_Z^2 + \sigma_X^2)(\sigma_Z^2+\sigma_Y^2)}}$


(b) Do things change?

No change. For $Cov(U,V)$, No matter its$Z+X$ and $Z+Y$ or $Z-X$ and $Z-Y$, Covariance of two different independent RVs will be eliminated. Therefore $Cov(U,V) = Var(X)$ here still.

$\rho_{U,V}$ will not change due to the same reason above.



(c) $\sigma_Z^2$ much larger than the other two?

$\rho_{U,V}=\frac{\sigma_Z^2}{\sqrt{(\sigma_Z^2 + \sigma_X^2)(\sigma_Z^2+\sigma_Y^2)}} \sim\frac{\sigma_Z^2}{\sqrt{(\sigma_Z^2 )(\sigma_Z^2)}} \sim 1$

therefore $\rho_{U,V}$ converges to 1

**he correlation will be dominated by Z.**

(d) $\sigma_Z^2$ much smaller than the other two?

$\rho_{U,V}$ converges to $\frac{\sigma_Z^2}{\sqrt{\sigma_X^2 \sigma_Y^2}} \sim 0$


(e):
generally, the formula for calculating the correlation coefficient standardizes the variable, and changes in scale or units of measurement will not affect its value. （quote from online
However, if the standard deviation of z becomes way larger and smaller than sd of x and y as shown in (c) and (d) above, in other words, if the dispersion of z is relatively too larger or smaller than the dispersion of x and y, Cor(U,V) here will be affected.

**The standardization of variables prior to statistical analysis (e.g., predictors in a regression model) ismotivated by similar logic, i.e., to prevent variables with large variances from dominating the solution. **

Problem 2 (Based on Rice 4.64):

(a):
$Cov(\bar{X}, \bar{Y}) = E[\bar{X}\bar{Y}] - E[\bar{X}]E[\bar{Y}] = E[\frac{X-\mu_X}{\sqrt{\sigma_X^2}} \frac{Y-\mu_Y}{\sqrt{\sigma_Y^2}}] - E[\frac{X-\mu_X}{\sqrt{\sigma_X^2}}] E[\frac{Y-\mu_Y}{\sqrt{\sigma_Y^2}}]$

$Cov(\bar{X}, \bar{Y})= \frac{1}{\sqrt{\sigma_X^2 \sigma_Y^2}} E[(X-\mu_X)(Y-\mu_Y)] - \frac{1}{\sqrt{\sigma_X^2 \sigma_Y^2}}E[X-\mu_X]E[Y-\mu_Y]$

$= \frac{1}{\sqrt{\sigma_X^2 \sigma_Y^2}} [E[XY-\mu_y X - \mu_X Y + \mu_X \mu_Y] - (E[X] - \mu_X)(E[Y] - \mu_Y)]$

$= \frac{1}{\sqrt{\sigma_X^2 \sigma_Y^2}} [E[XY] - E[X]E[Y]] = \rho_{X,Y}$

therefore $Cov(\bar{X}, \bar{Y}) = \rho_{X,Y}$



(b):
```{r}
?prcomp()
# set a 4 x 4 matrix
V1 <- c(5,10,15,20)
V2 <- seq(7:10)
V3 <- c(6,11,9,28)
V4 <- c(38:35)
matrix <- cbind(V1,V2,V3,V4)
covariance <-cov(matrix)

#PCA: eigenvalue decomposition of the sample covariance matrix
prcomp(covariance, scale = T, center = F)
# tried scale = T or F and center = T or F, noted that the defalt is scale = F while center = T.

# if center = T, the variables will be shifted to be zero centered.
# if scale = T, the variables are scaled to have unit variance be fore the analysis takes place

# we set scale = T and center = T when we want to get a scaled pca on the data matrix 
# which centers at 0, especially when we don't have the scaling done before doing pca. 

# it is good to scale the data in most cases, otherwise the magnitude to the variables 
# dominates the associations between the variables. Unless all the variables are previously 
# recorded in the same scale, it is better to normalize (scale) the data. For example, 
# do pca on a gene matrix.

# similarly, better not to have the uncentered matrix data.
```


Problem 3 (Based on Rice 4.74):

(a) expected number of offspring in the third generation:
Let $E[Generation 1] = E[G_1] = \mu$
and $Var[G_1] = \sigma^2$

$E[G_2|G_1] = G_1 \mu$, $G_1$ is the population of the first generation
$E[G_3 | G_2] = G_2 \mu$

the second generation is: $E[E[G_2|G_1]] = E[G_2] = E[G_1 \mu] = \mu E[G_1] = \mu^2$
and the third generation is: $E[E[G_3|G_2]] = E[G_3] = E[G_2 \mu] = \mu E[G_2] = \mu^3$

** $E[T3] = \mu^2$  **

(b) ariance of the number of offspring in the third generation:

$Var[G_1] = \sigma^2$, 

$Var[G_2|G_1] = G_1 \sigma^2$

$Var[G_3|G_2] = G_2 \sigma^2$

the second generation is: $E[Var[G_2|G_1]] = Var[G_2] = Var[G_1 \sigma^2] = \sigma^2 E[G_1] = \sigma^2 \mu$
and the third generation is: $E[Var[G_3|G_2]] = Var[G_3] = Var[G_2 \sigma^2] = \sigma^2 E[G_2] = \sigma^2 \mu^2$

**$Var(T3) = \mu \sigma^2 (\mu +1)$**

(c)
```{r}

# generate 1000 populations
population <-matrix(0, 1000, 1)
for (i in 1:nrow(population)){
m=1
while(m<1001){
G2<-rpois(1,2) # set first generation with 1 person, plug the poisson distribution with lembda = 2 to the second generation and solve for # of people in G2
G3<-rpois(G2,2) # the second generation has G2 people. Solve for G3 from each individual in G2
G<-sum(G3) # add each G3 to get the overall G3 in per population (picture the population tree)
population[i,1]<-G
m=m+1
}
}

hist(population[,1], main = "Number of offsprings in the third generation")

mean(population)
var(population)
```


Problem 5 (Based on Rice 5.1): need to edit

given that:
$E[X_i] = \mu$, and $Var(X_i) = \sigma_i$
and $\frac{1}{n^2}[\sigma_1^2 + \sigma_2^2 + ... + \sigma_n^2] \to 0$

so $Var(\bar{X_n}) = \frac{1}{n^2} \sum Var(X_i) = \frac{1}{n^2} (\sigma_1^2 + \sigma_2^2 + ...+ \sigma_n^2) \to 0$

According to Chebyshev's inequality, 
$P(| \bar{X_n} - \mu| > \epsilon) \leqslant \frac{Var(\bar{X_n})}{\epsilon^2}  \to 0$ since $Var(\bar{X_n}) \to 0$


Problem 7 (Based on Rice 5.16):

(a) approximate $P(S \leqslant10)$

given $S = X_1 + X_2 +...+X_{20}$, $f(x) = 2x$ when $0 \leqslant x \leqslant 1$

$E[X_i] = \int_{0}^{1} x f(x) dx = \int_{0}^{1} 2x^2 dx = \frac{2}{3}$

$E[X_i]^2 = \int_{0}^{1}x^2 f(x)dx =\int_{0}^{1}2x^3 dx = \frac{1}{2}$
$Var(X_i) = E[X^2] - E[X]^2 = \frac{1}{2} - (\frac{2}{3})^2 = \frac{1}{18}$

According to CLT:
$P(S \leqslant 10) \sim \lim_{n\to\infty} P(\frac{S_n}{\sqrt{n} \sigma} \leqslant x)$ for $x = {1,2,3,...,20}$

then $P(S \leqslant 10) \sim P(\frac{S-\frac{40}{3}}{\sqrt{\frac{20}{18}}} \leqslant \frac{10-\frac{40}{3}}{\sqrt{\frac{20}{18}}}) = P(\frac{S-\frac{40}{3}}{\sqrt{\frac{20}{18}}} \leqslant -\sqrt{10})$ 

$P(S \leqslant 10) \sim \phi(-\sqrt{10}) = 0.00079$


(b):


library(dplyr)

x <- seq(0,20, length =1000)

# mean = n times mu = 20*2/3 = 13.333, where n = 20
# sd = sigma * sqrt(n) = sqrt(20/18) = 1.054
y <- dnorm(x, mean=13.333, sd=1.054) # CLT-based density

mati <- (runif(20000,0,1)/2)

mati2 <- matrix(mati, byrow=TRUE, nrow=10000, ncol = 20 )
mati2 <- as.data.frame(mati2)
mati3 <- mati2 %>% 
mutate(rowsums= rowSums(mati2))

# plot true density
plot(density(mati3$rowsums), ylim=c(0,0.8),xlim=c(0,20), main = "density of S")

# plot CLT based density
lines(x, y, type="l", lwd=1, col = 'blue')

# they look off-shifted. It is because CLT is to approximate
# large-sample-sized variables. Here we only have 20 variables.

***
```{r}

 n = 10000
sim.vals = matrix(sqrt(runif(n*20)), nrow=n)
sum.vals = apply(sim.vals, 1, sum)
plot(density(sum.vals), xlab="x", ylab="f(x)", xlim=c(0,20))
x.vals = seq(from=0,to=20, by=0.01)
points(x.vals, dnorm(x.vals, mean=40/3, sd=0.236*sqrt(20)), type="line", col="red")
```


***



Problem 8 (Based on Rice 5.21):

(a):

$\hat{I}(f) = \frac{1}{n} \sum f(X_i)/ g(X_i)$ on $[a,b]$

$E[\hat{I}(f)] = \int_{a}^{b} \hat{I}(f) g(X_i)dx = \int_{a}^{b} \frac{1}{n}\sum \frac{f(X_i)}{g(X_i)} g(X_i) dx = \int_{a}^{b} \frac{1}{n} \sum f(X_i) dx = \int_{a}^{b} f(x) dx = I(f)$


(b):


mean <- 0
sd <- 1

n <- seq(5,100) # generate an array of n so I can plot the n and i hat thing later
for (m in n){
   I_hat <- 0
   for (i in 1:m){
    fx <- dnorm(i, mean, sd)  # generate the density of standard normal           distribution for each n
    gx <- 1 # generate the density of the standard uniform distribution for each n
     I_hat <- fx / gx + I_hat # this is actually partial of I_hat without multiplied by 1/n which i will do it outside of this inner loop
  }
   I_hat <- (1/n) * I_hat
}

plot(n, I_hat, main = "I_hat as a function of n", xlab = "n", ylab = "I_hat")
weighted.mean(I_hat)
#E(I_hat) = 0.0097
abline(h = 0.0097, col = 'purple')

# now lets plot I(f)
# integral

x <- seq(0,1)
f <- function(x) dnorm(x, mean=0, sd=1)
I_f <- integrate(f, lower=0, upper=1)
# I_f = 0.3413477
abline(h= 0.3413477, col = 'pink')



***
```{r}
(I_f = pnorm(1) - pnorm(0))

n.vals = 5:100
x.vals = runif(100)
g.vals = rep(1,100) # U(0,1) density is 1 in region
f.vals = dnorm(x.vals)
I.est = rep(0, length(n.vals))
for (i in 1:length(n.vals)) {
  n = n.vals[i]
  I.est[i] = mean(f.vals[1:n]/g.vals[1:n])
  }
I.est[1:10]

```

***


