---
title: "QBS177_HW3"
output: html_document
---

# QBS 177 HW 3

### 1.
```{r}
PCA.1 <- read.table("~/Documents/QBS 177/PCA.example1.txt", header=TRUE)

# Fill the missing values using column means.
for(i in 1:ncol(PCA.1)){
PCA.1[is.na(PCA.1[,i]), i] <- round(mean(PCA.1[,i], na.rm = TRUE))
}
sum(is.na(PCA.1))

date()
PCA.1.pc.eigen.cor <- prcomp(PCA.1)
date()

# Check the output object and identify the principle components.
names(PCA.1.pc.eigen.cor)
PCA1.scores.eigen <- PCA.1.pc.eigen.cor$x
```

### 2.
```{r}
PCA1.1 <- scale(PCA.1, scale = FALSE)
date()
PCA1.pc.scale.svd <- svd(PCA1.1)
date()

# 12.	Check the output object and compare with decomposition result
names(PCA1.pc.scale.svd)
head(PCA1.pc.scale.svd$d^2/(nrow(PCA.1)-1))
head(PCA.1.pc.eigen.cor$sdev^2)

PCA.1.pc.eigen.cor$rot[1:5,1:5]
PCA1.pc.scale.svd$v[1:5,1:5]
sum(abs(PCA1.pc.scale.svd$v-PCA.1.pc.eigen.cor$rot))

```
The result with spectral decomposition is the same compared to singular value decomposition while singular value decomposition runs less time than spectral decomposition.


### 3.
```{r}
(PCA1.euro <- apply(PCA1.scores.eigen[19662:19826,1:2], 2, mean))
(PCA1.asian <- apply(PCA1.scores.eigen[19827:19963,1:2], 2, mean))
(PCA1.african <- apply(PCA1.scores.eigen[19964:20166,1:2],2,mean))

(PCA1.unknown <- apply(PCA1.scores.eigen[1:19661,1:2],2,mean))

PCA1.unknown-PCA1.euro
PCA1.unknown-PCA1.asian
PCA1.unknown-PCA1.african

```
After comparing the unknown population's centroid and comparing to the other three populations, we find out that the unknown population's result is closest to the European population.




### 4. 
```{r}
# Visualize the first two principle components.
plot(PCA1.scores.eigen[,1], PCA1.scores.eigen[,2])

# replot with color coding
plot(PCA1.scores.eigen[1:19661,1],PCA1.scores.eigen[1:19661,2],type="p",col="black",pch=1,xlab="PC1", ylab="PC2", xlim = range(PCA1.scores.eigen[,1]), ylim = range(PCA1.scores.eigen[,2]));
points(PCA1.scores.eigen[19662:19826,1],PCA1.scores.eigen[19662:19826,2],col="red",pch=20); # euro
points(PCA1.scores.eigen[19827:19963,1],PCA1.scores.eigen[19827:19963,2],col="green",pch=20) # asian
points(PCA1.scores.eigen[19964:20166,1],PCA1.scores.eigen[19964:20166,2],col="blue",pch=20) # african
title(main="Principal Components Analysis using princomp in R", col.main="black", font.main=1,outer=T)

```