---
title: "QBS121 Assignment 7"
output: html_document
---

# Assignment on Propensity Scores and Missing Data

## 1 Mising Data


```{r}
n = 10^5
X = runif(n) < 0.5
Z = rnorm(n)
Noise = rnorm(n)
Y = X + Z + Noise

summary(lm(Y ~ X + Z))

Zm = ifelse(runif(n) < 0.40, NA, Z)
DF = data.frame(Y, X, Zm)
```

### 1.   Is Zm missing at random?
Zm is missing at random. Zm = ifelse(runif(n) < 0.40, NA, Z) means if runif(n)<0.4, then Zm is NA; if runif(n) >= 0.4, then Zm = Z. While runif is to generate random deviates with equally possible values, so Zm basically just generates an array where 40% of them are missing values.


### 2. Impute Zm by its mean
```{r}
Zi <- ifelse(is.na(Zm), mean(Zm, na.rm=TRUE), Zm)
summary(lm(Y ~ X + Zi))$coef
```


### 3. 
From the results above, we can see that the coefficiens from the imputed Zm model is relatively different from the true values from the unimputed Zm model.


### 4.
```{r}
summary(lm(Y ~ X + Z))$r.squared
summary(lm(Y ~ X + Zi))$r.squared

```
The r-squared is relatively different from the model in which Zi was used.


### 5.
Use imputation formed by other covariates:
```{r}
Z.lm <- lm(Y ~ X + Z)
Z.fit <- ifelse(is.na(Z), cbind(1,X) %*% Z.lm$coef, Z)
summary(o.mn.imp <- lm(Y ~ X+Z.fit))

```
Now the coefficients and R-squared value from the reimputed Z model are very close to the original Z model.



## 2 ropensity Scores

### 1. Odds ratio
```{r}
setwd("~/Documents/QBS 121")
ho <-  read.delim("~/Documents/QBS 121/Teaching Hospital Outcomes.txt")
ho <- as.data.frame(ho)
os <- summary(glm(TeachingHospital ~ ., data=ho, family=binomial))
exp(os$coef["UNFAVDX",1])
```


### 2. Propensity
```{r}
Model <- (o <- glm(UNFAVDX ~ (Age+Female+Black+Race.Other+Hispanic+DM+HTN)^2,data=ho, family="binomial"))
Propensity <- Model$fit
boxplot(split(Propensity,ho$UNFAVDX))
IWP <- ifelse(ho$UNFAVDX,1/Propensity, 1/(1 - Propensity))

```


### 3. 10 bins of the propensity score:
```{r}
library(randomForest)
Bins <- cut(Propensity,quantile(Propensity, (0:10)/10), include.lowest = TRUE)
(os1 <- summary(glm(ho$TeachingHospital ~ ho$UNFAVDX + Bins, family=binomial)))$coef[2,]
exp(os1$coef[2,1])

```


### 4.  Calculate inverse propensity weights
```{r}
(os2 <- summary(glm(ho$TeachingHospital ~ ho$UNFAVDX, weight=IWP)))$coef[2,]
exp(os2$coef[2,1])
```